---
title: "R Notebook"
output: html_notebook
---

```{r}
orders %>%
    mutate(offspring.year = litter.size * litters.year) %>% 
    ggplot(aes(x=offspring.year, y=weaning.mo.)) +
               geom_point() +
    geom_smooth(method='loess', alpha=0.5)
```
#correlation between offspring/year and weaning time

```{r}
orders %>%
    ggplot(aes(x=gestation.mo., y=weaning.mo.)) +
    geom_point() +
    geom_smooth(method='loess', alpha=0.5)
```
#correlation between gestation time and weaning time

```{r}
orders %>%
    ggplot(aes(x=AFR.mo., y=max..life.mo.)) +
    geom_point() +
    geom_smooth(method='loess', alpha=0.5)
```
#correlation between age of first reproduction and adult lifespan

```{r}
orders %>%
    mutate(offspring.year = litter.size * litters.year) %>%
    ggplot(aes(x=AFR.mo., y=offspring.year)) +
    geom_point() +
    geom_smooth(method='loess', alpha=0.5)
```
#correlation between age of first reproduction and offspring/year

```{r}
orders %>%
    ggplot(aes(x=max..life.mo., y=AFR.mo.)) +
    geom_point() +
    geom_smooth(method='loess', alpha=0.5)
```
#correlation between max lifespan and age of first reproduction

```{r}
orders %>%
    mutate(offspring.year = litter.size * litters.year) %>%
    ggplot(aes(y=max..life.mo., x=offspring.year)) +
    geom_point()
```
#correlation between offspring/year and max lifespan

